---
title: "Sequential trials"
output: html_document
date: '2022-08-17'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(rjags)
library(HDInterval)
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)

library(patchwork)
library(stringr)
library(pbapply)
library(parallel)

library(ggplot2)
library(extraDistr)
source('./R/prior.data.func.R')
source('./R/call_jags.R')
source('./R/sim.data.R')
source('./R/jags_basic_pois.R')
source('./R/jags_modified_pois_commensurate_gamma.R') #NOTE, this is experimental version

source('./R/caterplot.func.R')
source('./R/prior_post_compare.R')
source('./R/extract_parm.R')

source('./R/run_all_models.R')

N.sim=500

```


VE from Novavax. This function takes the mean and 95% CI of vaccine effectiveness, converts to a log(RR), and estimates the precision fro the CIs, assuming it is symmetric around the mean (on scale of log(RR))
```{r}
#This is main novavax result
#prior.data <- prior.data.func(ve.obs.mean=39.4, ve.obs.lcl=5.3, ve.obs.ucl=61.2, N_cases_orig =c(35,41), pop_orig= c(1430, 2765))

#Table 2: per protocol 90 day VE against RSV LRTI with hospitalization
prior.data <- prior.data.func(ve.obs.mean=44.4, ve.obs.lcl=19.6, ve.obs.ucl=61.5, N_cases_orig =c(53,57), pop_orig= c(1430, 2765))
```

Set parameters for new trial population

```{r}


#proportion of kids in placebo group who get the outcome set to 0.037 for Noavavx secondary endpoint
set.rate.control <- 0.037 

#VE in the new trial, if it agrees with original
set.ve.new.trial.agree = prior.data$ve.obs.mean

#discrepant VE in new trial (higher)
set.ve.new.trial.higher = 90

```

## Scenario 1: Same VE as original trial
Novavax had 3045 vaccinees and 1581 controls; 2.4% in placebo group had primary endpoint in 90 days
```{r}
set.seed(123)

sim1 <- sim.data.func(N.vax = 10000,
                     N.control = 10000,
                     rate.control = set.rate.control,
                     ve.new.trial=set.ve.new.trial.agree,
                     n.sim=N.sim)
  
```

## Scenario 2: no actual effect
```{r}
set.seed(456)

sim2 <- sim.data.func(N.vax = 10000,
                     N.control = 10000,
                     rate.control = set.rate.control,
                     ve.new.trial=0,
                     n.sim=N.sim)
```

## Scenario 3: discrepant actual effect
```{r}
set.seed(456)

sim3 <- sim.data.func(N.vax = 10000,
                     N.control = 10000,
                     rate.control = set.rate.control,
                     ve.new.trial=set.ve.new.trial.higher,
                     n.sim=N.sim)
```


## At which points do we want to evaluate VE?

How many vaccinees are enrolled?

```{r}

eval_points = seq(from=100, to=2000, by=100)

```

test run. 500 vaccinees+500 controls. It takes 0.27 seconds for 1 run. Need to test 20 cut points, 3 true VE levels, and 500 sims (30,000 combos). This would take 135 minutes if run sequentially. 


generate combined dataset for modeling--need to fill 0s if stratum not represented
```{r}
mod.ds1 <- lapply(eval_points , function(set.pop){
    ds1 <- bind_rows(sim1, sim2, sim3) %>%
        group_by(repN,ve.new.trial, treatment) %>%
      mutate(pop=set.pop) %>%
      filter( ((ID<pop &treatment=='placebo')|(ID<pop &treatment=='vax') )) %>%
          summarize(N_cases= n(), pop=mean(pop)) %>%
      ungroup() 
    
return(ds1)
    }) %>%
  bind_rows() %>%
  tidyr::complete(repN,ve.new.trial, treatment , pop,   fill=list(N_cases=0)) %>%
    mutate( vax=if_else(treatment=='vax',1,0)) 

```

Accrual of cases in the 500 trials
```{r}
mod.ds1 %>%
  filter(ve.new.trial==44.4) %>%
  ggplot(aes(x=pop, y=N_cases, group=interaction(repN, treatment), color=treatment))+
  geom_line(alpha=0.1) +
  theme_classic()
```





```{r}
ptm <- proc.time()

mod1 <- mod.ds1 %>%
  filter( repN==2  & pop==500 &  ve.new.trial==44.4 ) %>%
  call_jags( prior.mean=0, prior.prec=1e-4)

proc.time() - ptm

```


Call all models for both simulated datasets. Use parallel processing
```{r, eval=F}
ptm <- proc.time()


library(multidplyr)
ptm <- proc.time()
numCores = detectCores() -1
cluster1 <- new_cluster(numCores)
cluster_library(cluster1, "dplyr")
cluster_library(cluster1, "rjags")
cluster_library(cluster1, "HDInterval")
cluster_copy(cluster1, "call_jags")
cluster_copy(cluster1, "prior.data")
cluster_copy(cluster1, "model_string_basic_pois")
cluster_copy(cluster1, "model_string_commensurate_gamma")

#baseline model without pooling
mod1 <- mod.ds1 %>%
  group_by(repN,pop,ve.new.trial) %>%
  multidplyr::partition(cluster=cluster1)  %>%
 do(call_jags(., prior.mean=0, prior.prec=1e-4)) %>%
  collect() %>%
  ungroup() %>%
  mutate(Pooling_type='none')

#full pooling
mod2 <- mod.ds1 %>%
  group_by(repN,pop,ve.new.trial) %>%
 # filter(rep<=3) %>%
  multidplyr::partition(cluster=cluster1)  %>%
 do(call_jags(., prior.mean=prior.data$log_irr.obs[1], prior.prec=prior.data$prec.log.irr.obs)) %>%
  collect() %>%
  ungroup() %>%
  mutate(Pooling_type='Full')

#Commensurate prior pooling
mod3 <- mod.ds1 %>%
  group_by(repN,pop,ve.new.trial) %>%
  multidplyr::partition(cluster=cluster1)  %>%
 do(call_jags(., prior.mean=prior.data$log_irr.obs[1], prior.prec=prior.data$prec.log.irr.obs, model.select=model_string_commensurate_gamma)) %>%
  collect() %>%
  ungroup() %>%
  mutate(Pooling_type='Commensurate')

proc.time() - ptm

all.mods <- bind_rows(mod1, mod2, mod3)

saveRDS(all.mods, './Results/all.mods_novavax.rds')

ptm <- proc.time()

```

```{r}
all.mods <- readRDS( './Results/all.mods_novavax.rds')

```

## Bayesian estimates of performance

```{r}

#the p_xx are replicated for each parameter so filter to just 1 parameter
probs <- all.mods %>% 
  group_by(repN, pop, ve.new.trial,Pooling_type) %>%
  filter(parm=='beta1')  %>%
  ungroup()
```

### Trajectories for 500 trials

This shows how the probabilities evolve as more data accrue. This shows probability of any benefit: P(VE>0|data)

```{r}
probs %>%
  filter(ve.new.trial==44.4 & Pooling_type != 'Full') %>%
  mutate(Pooling_type = as.factor(Pooling_type)) %>%
ggplot( aes(x=pop, y=p_any_benefit0, group=repN, color=Pooling_type)) +
  geom_line( alpha=0.1) +
  theme_classic() +
  facet_wrap(~Pooling_type)
```

This shows how the probabilities evolve as more data accrue. This shows probability of any benefit: P(VE>25|data)


```{r}
probs %>%
  filter(ve.new.trial==44.4 & Pooling_type != 'Full') %>%
  mutate(Pooling_type = as.factor(Pooling_type)) %>%
ggplot( aes(x=pop, y=p_moderate_0_75, group=repN, color=Pooling_type)) +
  geom_line( alpha=0.1) +
  theme_classic() +
  facet_wrap(~Pooling_type)
```

### Summaries of the trajectories 

What is the probability of detecting any effect if there is (A) No effect (B) a moderate effect in line with the original trial or (C) A large effect, bigger than original trial? The shaded region shows where 90% of the estimates lie

```{r}
probs %>%
  filter(Pooling_type != 'Full') %>%
  mutate(Pooling_type = as.factor(Pooling_type), ve.new.trial=as.factor(ve.new.trial)) %>%
  group_by(pop, ve.new.trial,Pooling_type) %>%
  summarize(p_any_benefit0_mean=mean(p_any_benefit0), p_any_benefit0_lcl=quantile(p_any_benefit0, probs=0.05),
             p_any_benefit0_ucl=quantile(p_any_benefit0, probs=0.95)) %>%
ggplot( aes(x=pop, y=p_any_benefit0_mean,  color=Pooling_type, group=Pooling_type, fill=Pooling_type)) +
  geom_line() +
  geom_ribbon( aes(x=pop, ymin=p_any_benefit0_lcl, 
                   ymax=p_any_benefit0_ucl,
                   color=Pooling_type, group=Pooling_type), alpha=0.1) +
  theme_classic() +
  facet_wrap(~ve.new.trial)+
  ggtitle('Prob any effect')
```


What is the probability of detecting NO effect (or effect between 20% decline or 25% increase) when the data truly have: (A) No effect (B) a moderate effect in line with the original trial or (C) A large effect, bigger than original trial? the shaded region shows where 90% of the estimates lie

```{r}
probs %>%
  filter(Pooling_type != 'Full') %>%
  mutate(Pooling_type = as.factor(Pooling_type), ve.new.trial=as.factor(ve.new.trial)) %>%
  group_by(pop, ve.new.trial,Pooling_type) %>%
  summarize(p_mean=mean(p_similar_0_8__1_25), p_lcl=quantile(p_similar_0_8__1_25, probs=0.05),
             p_ucl=quantile(p_similar_0_8__1_25, probs=0.95)) %>%
ggplot( aes(x=pop, y=p_mean,  color=Pooling_type, group=Pooling_type, fill=Pooling_type)) +
  geom_line() +
  geom_ribbon( aes(x=pop, ymin=p_lcl, 
                   ymax=p_ucl,
                   color=Pooling_type, group=Pooling_type), alpha=0.1) +
  theme_classic() +
  facet_wrap(~ve.new.trial)+
  ggtitle('Prob no effect')
```


What is the probability of detecting a moderate effect or greater (at least 25% decline) effect when the data truly have: (A) No effect (B) a moderate effect in line with the original trial or (C) A large effect, bigger than original trial? the shaded region shows where 90% of the estimates lie

```{r}
probs %>%
  filter(Pooling_type != 'Full') %>%
  mutate(Pooling_type = as.factor(Pooling_type), ve.new.trial=as.factor(ve.new.trial)) %>%
  group_by(pop, ve.new.trial,Pooling_type) %>%
  summarize(p_mean=mean(p_moderate_0_75), p_lcl=quantile(p_moderate_0_75, probs=0.05),
             p_ucl=quantile(p_moderate_0_75, probs=0.95)) %>%
ggplot( aes(x=pop, y=p_mean,  color=Pooling_type, group=Pooling_type, fill=Pooling_type)) +
  geom_line() +
  geom_ribbon( aes(x=pop, ymin=p_lcl, 
                   ymax=p_ucl,
                   color=Pooling_type, group=Pooling_type), alpha=0.1) +
  theme_classic() +
  facet_wrap(~ve.new.trial)+
  ggtitle('Prob Moderate effect')
```

What is the probability of detecting a harmful effect or greater (an increase in disease rates) when the data truly have: (A) No effect (B) a moderate effect in line with the original trial or (C) A large effect, bigger than original trial? the shaded region shows where 90% of the estimates lie

```{r}
probs %>%
  filter(Pooling_type != 'Full') %>%
  mutate(Pooling_type = as.factor(Pooling_type), ve.new.trial=as.factor(ve.new.trial)) %>%
  group_by(pop, ve.new.trial,Pooling_type) %>%
  summarize(p_mean=mean(p_harm_o1), p_lcl=quantile(p_harm_o1, probs=0.05),
             p_ucl=quantile(p_harm_o1, probs=0.95)) %>%
ggplot( aes(x=pop, y=p_mean,  color=Pooling_type, group=Pooling_type, fill=Pooling_type)) +
  geom_line() +
  geom_ribbon( aes(x=pop, ymin=p_lcl, 
                   ymax=p_ucl,
                   color=Pooling_type, group=Pooling_type), alpha=0.1) +
  theme_classic() +
  facet_wrap(~ve.new.trial)+
  ggtitle('Prob Harmful effect')
```

Evolution of mean  estimates for log(VE)..suggests maybe things aren't convverging at low end
```{r}
probs %>%
  filter(ve.new.trial==44.4 & Pooling_type != 'Full') %>%
  mutate(Pooling_type = as.factor(Pooling_type)) %>%
ggplot( aes(x=pop, y=mean, group=repN, color=Pooling_type)) +
  geom_line( alpha=0.1) +
  theme_classic() +
  facet_wrap(~Pooling_type)+
  ylim(-5,5)
```


## Traditional frequentist measures of power, type 1 error

How do the three models perform when there is NO real effect?

1: no pooling 2: highly informative prior (full use) 3:Gamma, (4:Half-Cauchy 5: mixture model)
```{r, fig.width=8, fig.height=3}

 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='none' & pop==300) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Full' & pop==300) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Commensurate' & pop==300) %>%   caterplot.func()

ggplot()


```
For a larger dataset
```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='none' & pop==1000) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Full' & pop==1000) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Commensurate' & pop==1000) %>%   caterplot.func()


```


```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='none' & pop==1500) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Full' & pop==1500) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Commensurate' & pop==1500) %>%   caterplot.func()


```

How do the three models perform when there is a real effect, consistent with the original?

```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='none' & pop==300) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='Full' & pop==300) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='Commensurate' & pop==300) %>%   caterplot.func()
```

```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='none' & pop==500) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='Full' & pop==500) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='Commensurate' & pop==500) %>%   caterplot.func()


```

```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='none' & pop==900) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='Full' & pop==900) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='Commensurate' & pop==900) %>%   caterplot.func()


```

```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='none' & pop==1000) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='Full' & pop==1000) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='Commensurate' & pop==1000) %>%   caterplot.func()


```


```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='none' & pop==1500) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='Full' & pop==1500) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==44.4 & Pooling_type=='Commensurate' & pop==1500) %>%   caterplot.func()


```

## What if there is a discrepant effect between prior and new trial
```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==90 & Pooling_type=='none' & pop==300) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==90 & Pooling_type=='Full' & pop==300) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==90 & Pooling_type=='Commensurate' & pop==300) %>%   caterplot.func()


 all.mods %>%   filter(ve.new.trial==90 & Pooling_type=='none' & pop==1000) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==90 & Pooling_type=='Full' & pop==1000) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==90 & Pooling_type=='Commensurate' & pop==1000) %>%   caterplot.func()

  all.mods %>%   filter(ve.new.trial==90 & Pooling_type=='none' & pop==1500) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==90 & Pooling_type=='Full' & pop==1500) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==90 & Pooling_type=='Commensurate' & pop==1500) %>%   caterplot.func()

```
