---
title: "Sequential trials"
output: html_document
date: '2022-08-17'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rjags)
library(HDInterval)
library(dplyr)
library(patchwork)
library(stringr)
library(pbapply)
library(parallel)

library(ggplot2)
library(extraDistr)
source('./R/prior.data.func.R')
source('./R/call_jags.R')
source('./R/sim.data.R')
source('./R/jags_basic_pois.R')
source('./R/jags_modified_pois_commensurate_gamma.R') #NOTE, this is experimental version

source('./R/caterplot.func.R')
source('./R/prior_post_compare.R')
source('./R/extract_parm.R')

source('./R/run_all_models.R')

N.sim=500

```


```{r}
hist(rgamma(1000,1,0.001))

#SD
sd <- 1/sqrt(rgamma(10000,1,0.1))

mean(1/sqrt(rgamma(10000,1,0.1)))
median(1/sqrt(rgamma(10000,1,0.1)))

hist(1/sqrt(rgamma(10000,0.01,0.01)), main='SD')

 
        sigma <- abs(rlst(n=10000,df=1, sigma=1/sqrt(50), mu=0) )
  	    prec <- 1/(sigma^2)
# # 	    
# # 	    mean(prec)
# library(extraDistr)
# prec2 <- rhcauchy(10000,sigma=0.2)
```


VE from Phase 2b of Matisse This function takes the mean and 95% CI of vaccine effectiveness, converts to a log(RR), and estimates the precision fro the CIs, assuming it is symmetric around the mean (on scale of log(RR))
```{r}
#This is main novavax result
#prior.data <- prior.data.func(ve.obs.mean=39.4, ve.obs.lcl=5.3, ve.obs.ucl=61.2, N_cases_orig =c(35,41), pop_orig= c(1430, 2765))

#This is Novavax LMIC data only
#Table S16: per protocol 90 VE against RSV LRTI with hospitalization
prior.data <- prior.data.func(ve.obs.mean=91.5, ve.obs.lcl=-5.6, ve.obs.ucl=99.8, N_cases_orig =c(3,1), pop_orig= c(103, 405))
```

Set parameters for new trial population

```{r}

#Sample sizes for vaccinee group
N.vax.test <- c(300,500,  700, 900, 1000,1500, 3700 )

#Sample sizes for control group--1:1 randomization
N.control.test <- round(N.vax.test)

#proportion of kids in placebo group who get the outcome set to 2.9% (0.029) for Phase2b pfizer
set.rate.control <- 0.029


#VE in the new trial
set.ve.new.trial.agree = prior.data$ve.obs.mean

#Let's test power to detect a 60% effect

set.ve.new.trial.lower = 50

```

## Scenario 1: Same VE as original trial
Novavax had 3045 vaccinees and 1581 controls; 2.4% in placebo group had primary endpoint in 90 days
```{r}
set.seed(123)

sim1 <- mapply(FUN=sim.data.func, 
       N.vax=N.vax.test, 
       N.control=N.control.test,  
       rate.control=set.rate.control, 
       ve.new.trial=set.ve.new.trial.agree,
       n.sim=N.sim, SIMPLIFY = F)
```

## Scenario 2: no actual effect
```{r}
set.seed(456)

sim2 <- mapply(FUN=sim.data.func, 
       N.vax=N.vax.test, 
       N.control=N.control.test,  
       rate.control=set.rate.control, 
       ve.new.trial=0,
       n.sim=N.sim, SIMPLIFY = F)
```

## Scenario 3: moderate actual effect
```{r}
set.seed(456)

sim3 <- mapply(FUN=sim.data.func, 
       N.vax=N.vax.test, 
       N.control=N.control.test,  
       rate.control=set.rate.control, 
       ve.new.trial=set.ve.new.trial.lower,
       n.sim=N.sim, SIMPLIFY = F)
```

Call all models for both simulated datasets. Use parallel processing
```{r, eval=F}

library(multidplyr)
ptm <- proc.time()
numCores = detectCores() -1
cluster1 <- new_cluster(numCores)
cluster_library(cluster1, "dplyr")
cluster_library(cluster1, "rjags")
cluster_library(cluster1, "HDInterval")
cluster_copy(cluster1, "call_jags")
cluster_copy(cluster1, "prior.data")
cluster_copy(cluster1, "model_string_basic_pois")
cluster_copy(cluster1, "model_string_commensurate_gamma")

#baseline model without pooling
mod1 <- bind_rows(c(sim1, sim2, sim3)) %>%
  group_by(rep,pop,ve.new.trial) %>%
 # filter(rep<=20) %>%
  multidplyr::partition(cluster=cluster1)  %>%
 do(call_jags(., prior.mean=0, prior.prec=1e-4)) %>%
  collect() %>%
  ungroup() %>%
  mutate(Pooling_type='none')

#full pooling
mod2 <- bind_rows(c(sim1, sim2, sim3)) %>%
  group_by(rep,pop,ve.new.trial) %>%
 # filter(rep<=3) %>%
  multidplyr::partition(cluster=cluster1)  %>%
 do(call_jags(., prior.mean=prior.data$log_irr.obs[1], prior.prec=prior.data$prec.log.irr.obs)) %>%
  collect() %>%
  ungroup() %>%
  mutate(Pooling_type='Full')

#Commensurate prior pooling
mod3 <- bind_rows(c(sim1, sim2, sim3)) %>%
  group_by(rep,pop,ve.new.trial) %>%
 # filter(rep<=3) %>%
  multidplyr::partition(cluster=cluster1)  %>%
 do(call_jags(., prior.mean=prior.data$log_irr.obs[1], prior.prec=prior.data$prec.log.irr.obs, model.select=model_string_commensurate_gamma)) %>%
  collect() %>%
  ungroup() %>%
  mutate(Pooling_type='Commensurate')

proc.time() - ptm

all.mods <- bind_rows(mod1, mod2, mod3)

saveRDS(all.mods, './Results/all.mods.rds')
```

```{r}
all.mods <- readRDS( './Results/all.mods.rds')

```


How do the three models perform when there is NO real effect?

1: no pooling 2: highly informative prior (full use) 3:Gamma, (4:Half-Cauchy 5: mixture model)
```{r, fig.width=8, fig.height=3}

 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='none' & pop==300) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Full' & pop==300) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Commensurate' & pop==300) %>%   caterplot.func()


```
For a larger dataset
```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='none' & pop==1000) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Full' & pop==1000) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Commensurate' & pop==1000) %>%   caterplot.func()


```
```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='none' & pop==3700) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Full' & pop==3700) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==0 & Pooling_type=='Commensurate' & pop==3700) %>%   caterplot.func()


```

How do the three models perform when there is a real effect, consistent with the original?

```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==91.5 & Pooling_type=='none' & pop==300) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==91.5 & Pooling_type=='Full' & pop==300) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==91.5 & Pooling_type=='Commensurate' & pop==300) %>%   caterplot.func()
```

```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==91.5 & Pooling_type=='none' & pop==3700) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==91.5 & Pooling_type=='Full' & pop==3700) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==91.5 & Pooling_type=='Commensurate' & pop==3700) %>%   caterplot.func()


```
## What if there is a discrepant effect between prior and new trial
```{r, fig.width=6, fig.height=3}
 all.mods %>%   filter(ve.new.trial==50 & Pooling_type=='none' & pop==700) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==50 & Pooling_type=='Full' & pop==700) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==50 & Pooling_type=='Commensurate' & pop==700) %>%   caterplot.func()


 all.mods %>%   filter(ve.new.trial==50 & Pooling_type=='none' & pop==1000) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==50 & Pooling_type=='Full' & pop==1000) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==50 & Pooling_type=='Commensurate' & pop==1000) %>%   caterplot.func()

  all.mods %>%   filter(ve.new.trial==50 & Pooling_type=='none' & pop==1500) %>%   caterplot.func() +
 all.mods %>%   filter(ve.new.trial==50 & Pooling_type=='Full' & pop==1500) %>%   caterplot.func() +
   all.mods %>%   filter(ve.new.trial==50 & Pooling_type=='Commensurate' & pop==1500) %>%   caterplot.func()

```

